
---
title: "`ImageArray`"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
package: "`r BiocStyle::pkg_ver('ImageArray')`"
author: 
  - name: Artur Manukyan
output:
  BiocStyle::html_document
vignette: |
  %\VignetteIndexEntry{ImageArray}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r load-libs}
library(ImageArray)
library(BiocFileCache)
library(RBioFormats)
library(EBImage)
library(ggplot2)
library(shiny)
```

# Introduction

The *ImageArray* package provides a **unified, memory-efficient** framework for working with both **pyramidal** and **non-pyramidal** images in R, leveraging the DelayedArray infrastructure.  
With ImageArray you can store large images on disk (HDF5 or Zarr), treat them as array-like objects, perform lazy/blocked operations (e.g., rotate, flip, crop) across all pyramid levels, and seamlessly integrate with standard R image workflows.

# Installation

You can install `r Biocpkg("ImageArray")` from Bioconductor with:

```{r install-BiocManager, eval = FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE)) {
    install.packages("BiocManager")
}
BiocManager::install("ImageArray")
```

# Why pyramidal images?

Image pyramids are multi-resolution representations: starting from a full resolution image, you generate a series of down-sampled versions (e.g., half resolution, quarter resolution, etc). These are common in microscopy, digital pathology and large-scale imaging because:

- They allow fast zooming and visualization (you don’t always need full resolution to inspect large images).  
- They support scale-aware analysis (you may operate on coarse levels when doing overview tasks, fine levels for detail).  
- Storing data in a pyramidal stack often reduces memory footprint and enables on-disk processing.

In our context, an ImageArray object may contain **multiple series** (each a resolution level) internally, and most operations are applied *consistently* across all series.

# Usage

## EBImage

The main usage of the `r Biocpkg("ImageArray")` is to deliver 
`r Biocpkg("DelayedArray")` operations for images or a list of images in 
Pyramidal format. You can also use
`magick-image` objects.

```{r read}
# make random EBImage image
f = system.file("images", "sample.png", package="EBImage")
img = readImage(f)

# create image array
dir.create(td <- tempfile())
output_h5 <- file.path(td, "h5test")
imgarray <- writeImageArray(img, 
                         format = "HDF5ImageArray", 
                         output = output_h5, 
                         replace = TRUE)
imgarray
```

```{r visualize_read}
# visualize 
bfa.raster <- as.raster(imgarray, max.pixel.size = 300)
plot(bfa.raster)
```

Operations such as rotate, flip, flop or negate will be 
conducted on all levels

```{r rotate}
imgarray_rotated <- rotate(imgarray, degrees = 90)
imgarray_rotated

# visualize 
bfa.raster <- as.raster(imgarray_rotated, max.pixel.size = 300)
plot(bfa.raster)
```

You can even crop images, either by using `crop` or `[` method.

```{r crop}
imgarray_cropped <- imgarray[300:500, 200:300]
imgarray_cropped

# visualize 
bfa.raster <- as.raster(imgarray_cropped, max.pixel.size = 300)
plot(bfa.raster)
```

```{r crop2}
imgarray_cropped <- crop(imgarray, ind = list(300:500, 200:300))
imgarray_cropped

# visualize 
bfa.raster <- as.raster(imgarray_cropped, max.pixel.size = 300)
plot(bfa.raster)
```

## magick images

```{r read3}
# make random EBImage image
f = system.file("images", "sample.png", package="EBImage")
img = magick::image_read(f)

# create image array
dir.create(td <- tempfile())
output_h5 <- file.path(td, "h5test")
imgarray <- writeImageArray(img, 
                         format = "HDF5ImageArray", 
                         output = output_h5, 
                         replace = TRUE)

bfa.raster <- as.raster(imgarray, max.pixel.size = 300)
plot(bfa.raster)
```

## OME.TIFF (Bio-Formats) images

You can create `ImageArray` objects from OME.TIFF files with already defined
layers. 

```{r ometiff}
ome.tiff.file <- system.file("extdata", "xy_12bit__plant.ome.tiff", 
                             package = "ImageArray")
imgarray <- createImageArray(ome.tiff.file, series = 1, resolution = 1:2)
imgarray
```

Depending on the requested pixel size, functions may return the desired
resolution from the pyramid.

```{r ometiffvis, out.width="50%"}
bfa.raster <- as.raster(imgarray)
plot(bfa.raster)
dim(bfa.raster)
```

You can use either `max.pixel.size` or `min.pixel.size` to control the output

```{r ometiffvis2, out.width="50%"}
bfa.raster <- as.raster(img, max.pixel.size = 300)
plot(bfa.raster)
dim(bfa.raster)
```

# Use cases

The pyramid scheme introduced by `ImageArray` objects can also be used to 
generate scalable plots and interactive Shiny application for visualizing 
large images. 

## Visualizing H&E images

For this example, we will use a large H&E image used by 10x Genomics, 
generated after a Xenium in Situ platform run. 

```{r he}
image_file <- paste(
    "https://cf.10xgenomics.com/samples/xenium/1.0.1",
    "Xenium_FFPE_Human_Breast_Cancer_Rep1", 
    "Xenium_FFPE_Human_Breast_Cancer_Rep1_he_image.ome.tif",
    sep="/")
library(BiocFileCache)
bfc <- BiocFileCache()
image_file <- bfcrpath(bfc, image_file)
```

Let us create an `ImageArray` object first with the image above. The ome.tiff 
file includes multiple resolutions that we can ask `ImageArray` 
object to include.

```{r he_imgarray}
imgarray <- createImageArray(image_file, 
                             series = 1, 
                             resolution = 1:6)
imgarray
```
You can visualize subsets of large images really quickly using `crop` and
`as.raster`. Again, here we use `max.pixel.size` to regularize the resolution
parsed from the pyramid; that is, as defined below, only resolutions with
max pixel size 800 would be returned.

```{r speedy_vis, out.width="70%"}
# crop image
imgarray_sub <- crop(imgarray, ind = list(16000:19000, 7000:10000))

# convert to raster
img_raster <- as.raster(imgarray_sub, max.pixel.size = 800)

# plot with ggplot
imgggplot <- ggplot2::ggplot(data.frame(x = 0, y = 0), 
                             ggplot2::aes_string("x", "y")) + 
  ggplot2::coord_fixed(expand = FALSE, 
                       xlim = c(0, dim(img_raster)[2]), 
                       ylim = c(0, dim(img_raster)[1])) + 
  ggplot2::annotation_raster(img_raster, 0, dim(img_raster)[2], 
                             dim(img_raster)[1], 0, interpolate = FALSE)
imgggplot
```

## Interactive Shiny Applications

Now let us create a shiny application where we can interactively define 
subsetting and visualize the image quickly. 

The Shiny application will react quickly to regardless of the size of the 
image subset. This is because `as.raster` determines the layer that has the
highest resolution without exceeding the threshold (`max.pixel.size`) 
automatically.

```{r speedy_vis_shiny}
if(interactive()){
  
  # variables
  dimimg <- dim(imgarray)
  max.pixel.size <- 800
  
  # Define UI
  ui <- fluidPage(
    sliderInput("x_slider", label = "X", min = 1, 
                max = dimimg[1], value = c(16000,19000)),
    sliderInput("y_slider", label = "Y", min = 1, 
                max = dimimg[2], value = c(7000,10000)),
    mainPanel(
      plotOutput(outputId = "scatterPlot")
    )
  )
  
  # Define server logic
  server <- function(input, output) {
    
    output$scatterPlot <- renderPlot({
      
      # crop image
      indx <- seq(input$x_slider[1], input$x_slider[2])
      indy <- seq(input$y_slider[1], input$y_slider[2])
      imgarray_sub <- crop(imgarray, ind = list(indx, indy))
      
      # convert to raster
      img_raster <- as.raster(imgarray_sub, max.pixel.size = max.pixel.size)
      
      # plot with ggplot
      imgggplot <- ggplot2::ggplot(data.frame(x = 0, y = 0), 
                                   ggplot2::aes_string("x", "y")) + 
        ggplot2::coord_fixed(expand = FALSE, 
                             xlim = c(0, dim(img_raster)[2]), 
                             ylim = c(0, dim(img_raster)[1])) + 
        ggplot2::annotation_raster(img_raster, 0, dim(img_raster)[2], 
                                   dim(img_raster)[1], 0, interpolate = FALSE)
      imgggplot
    })
  }
  
  # Run the app
  shinyApp(ui = ui, server = server)
}
```

# Session info

```{r sessionInfo, echo=FALSE}
sessionInfo()
```
