
---
title: "`ImageArray`"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
package: "`r BiocStyle::pkg_ver('ImageArray')`"
author: 
  - name: Artur Manukyan
output:
  BiocStyle::html_document
vignette: |
  %\VignetteIndexEntry{ImageArray}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r load-libs}
library(ImageArray)
library(BiocFileCache)
library(RBioFormats)
library(EBImage)
library(ggplot2)
library(shiny)
```

# Introduction

The `r Biocpkg("ImageArray")` package provides a **unified, memory-efficient** 
framework for working with both **pyramidal** and **non-pyramidal** images in R 
by leveraging the `r Biocpkg("DelayedArray")` infrastructure. With 
`r Biocpkg("ImageArray")` you can store large images on disk (HDF5 or Zarr), 
treat them as array-like objects, perform delayed/lazy operations (e.g., 
rotate, flip, crop) across all pyramid levels, and seamlessly integrate with 
standard R image workflows.

# Installation

You can install `r Biocpkg("ImageArray")` from Bioconductor with:

```{r install-BiocManager, eval = FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE)) {
    install.packages("BiocManager")
}
BiocManager::install("ImageArray")
```

# Why pyramidal images?

Image pyramids are multi-resolution representations: starting from a full 
resolution image, you generate a series of down-sampled versions (e.g., half 
resolutions). These are common in microscopy, digital pathology and large-scale 
imaging because:

- They allow fast zooming and visualization (you don’t always need full 
resolution to inspect large images).  
- They support scale-aware analysis (you may operate on coarse levels when 
doing overview tasks, fine levels for detail).  
- Storing data in a pyramidal stack often reduces memory footprint and 
enables on-disk processing.

In our context, an ImageArray object may contain **multiple series** 
(each a resolution level) internally, and most operations are applied 
lazily across all series.

# Usage

## EBImage

We primarily use images via the `r Biocpkg("EBImage")` package. 

```{r read}
# make random EBImage image
f = system.file("images", "sample.png", package="EBImage")
img = readImage(f)
```

Similar to its counterparts (see `writeHDF5Array` in `r Biocpkg("HDF5Array")`), 
we use the `writeImageArray` function to write pyramid images to preferred 
locations. Here, `n.levels` control the amount of levels that the pyramid image
should have. 

```{r write}
# create image array
dir.create(td <- tempfile())
output_h5 <- file.path(td, "h5test")
imgarray <- writeImageArray(img, 
                            format = "HDF5ImageArray", 
                            output = output_h5, 
                            n.levels = 2,
                            replace = TRUE)
imgarray
```

Each level of a pyramid can be rasterized at any time.

```{r visualize_read}
# visualize 
bfa.raster <- as.raster(imgarray, level = 2)
plot(bfa.raster)
```

However, the true functionality of `r Biocpkg("ImageArray")` is to pick
pyramid levels in a memory-efficient level. Thus, packages and applications
that use `r Biocpkg("ImageArray")` can rasterize images by defining thresholds
for the pixel dimensions of the levels that are realized to the memory.

Here, by using the `max.pixel.size`, we request `r Biocpkg("ImageArray")` to 
return a pyramid level whose both width (`X`) and height (`Y`) are lower than
300. 

```{r visualize_read2}
# visualize 
bfa.raster <- as.raster(imgarray, max.pixel.size = 300)
dim(bfa.raster)
```

Operations such as rotate, flip, flop or negate will be 
conducted on all levels (without loading the image on the memory)
which then can be rasterized and plotted.

```{r rotate}
imgarray_rotated <- rotate(imgarray, degrees = 90)
imgarray_rotated

# visualize 
bfa.raster <- as.raster(imgarray_rotated, max.pixel.size = 300)
plot(bfa.raster)
```

You can even crop images, either by using `crop` or `[` methods.

```{r crop}
imgarray_cropped <- imgarray[300:500, 200:300]
imgarray_cropped

# visualize 
bfa.raster <- as.raster(imgarray_cropped, max.pixel.size = 300)
plot(bfa.raster)

imgarray_cropped <- crop(imgarray, ind = list(300:500, 200:300))
imgarray_cropped

# visualize 
bfa.raster <- as.raster(imgarray_cropped, max.pixel.size = 300)
plot(bfa.raster)
```

## magick images

`r Biocpkg("ImageArray")` is also compatible with the `r CRANpkg("magick")` 
package where we retain the dimensionality of the image similar to 
`r Biocpkg("EBImage")`.

```{r read3}
# make random EBImage image
f = system.file("images", "sample.png", package="EBImage")
img = magick::image_read(f)

# create image array
dir.create(td <- tempfile())
output_h5 <- file.path(td, "h5test")
imgarray <- writeImageArray(img, 
                         format = "HDF5ImageArray", 
                         output = output_h5, 
                         replace = TRUE)

# plot
bfa.raster <- as.raster(imgarray, level = 2)
plot(bfa.raster)
```

## OME.TIFF (Bio-Formats) images

You can create `ImageArray` objects from OME-TIFF files (or image formats that 
are compatible with Bio-formats, see `r Biocpkg("RBioFormats")`) with 
already defined layers. 

```{r ometiff}
ome.tiff.file <- system.file("extdata", "xy_12bit__plant.ome.tiff", 
                             package = "ImageArray")
RBioFormats::read.metadata(ome.tiff.file)

# define ImageArray object
imgarray <- createImageArray(ome.tiff.file, series = 1, resolution = 1:2)
imgarray
```

You can again use either `max.pixel.size` or `min.pixel.size` to control 
the output or level.

```{r ometiffvis2, out.width="50%"}
bfa.raster <- as.raster(img, max.pixel.size = 300)
plot(bfa.raster)
dim(bfa.raster)

bfa.raster <- as.raster(img, level = 1)
plot(bfa.raster)
dim(bfa.raster)
```

# Use cases

The delayed pyramid scheme introduced by `ImageArray` objects can also 
be used to generate scalable plots and interactive Shiny application 
for visualizing large images without loading them into the memory. 

## Visualizing H&E images

For this example, we will use a large H&E image used by 10x Genomics, 
generated after a Xenium in Situ platform run (i.e. postXenium H&E).

```{r he}
image_file <- paste(
    "https://cf.10xgenomics.com/samples/xenium/1.0.1",
    "Xenium_FFPE_Human_Breast_Cancer_Rep1", 
    "Xenium_FFPE_Human_Breast_Cancer_Rep1_he_image.ome.tif",
    sep="/")
library(BiocFileCache)
bfc <- BiocFileCache()
image_file <- bfcrpath(bfc, image_file)
```

Let us create an `ImageArray` object first with the image above. The OME-TIFF 
file includes multiple resolutions that we can ask `ImageArray` 
object to include.

```{r he_imgarray}
imgarray <- createImageArray(image_file, 
                             series = 1, 
                             resolution = 1:6)
imgarray
```

You can visualize subsets of large images really quickly using `crop` and
`as.raster`. Again, here we use `max.pixel.size` to regularize the resolution
parsed from the pyramid; that is, as defined below, only resolutions whose 
width and height are both under size 800 would be returned.

```{r speedy_vis, out.width="70%"}
# crop image
imgarray_sub <- crop(imgarray, ind = list(16000:19000, 7000:10000))

# convert to raster
img_raster <- as.raster(imgarray_sub, max.pixel.size = 800)

# plot with ggplot
ggplot2::ggplot(data.frame(x = 0, y = 0), 
                             ggplot2::aes_string("x", "y")) + 
  ggplot2::coord_fixed(expand = FALSE, 
                       xlim = c(0, dim(img_raster)[2]), 
                       ylim = c(0, dim(img_raster)[1])) + 
  ggplot2::annotation_raster(img_raster, 0, dim(img_raster)[2], 
                             dim(img_raster)[1], 0, interpolate = FALSE)
```

## Interactive Shiny Applications

Now let us create a shiny application where we can interactively define 
subsetting and visualize the OME-TIFF image quickly. 

The Shiny application will react quickly to regardless of the size of the 
image subset. This is because `as.raster` determines the layer that has the
highest resolution without exceeding the threshold (`max.pixel.size`) 
automatically.

```{r speedy_vis_shiny}
if(interactive()){
  
  # variables
  dimimg <- dim(imgarray)
  max.pixel.size <- 800
  
  # Define UI
  ui <- fluidPage(
    sliderInput("x_slider", label = "X", min = 1, 
                max = dimimg[1], value = c(16000,19000)),
    sliderInput("y_slider", label = "Y", min = 1, 
                max = dimimg[2], value = c(7000,10000)),
    mainPanel(
      plotOutput(outputId = "scatterPlot")
    )
  )
  
  # Define server logic
  server <- function(input, output) {
    
    output$scatterPlot <- renderPlot({
      
      # crop image
      indx <- seq(input$x_slider[1], input$x_slider[2])
      indy <- seq(input$y_slider[1], input$y_slider[2])
      imgarray_sub <- crop(imgarray, ind = list(indx, indy))
      
      # convert to raster
      img_raster <- as.raster(imgarray_sub, max.pixel.size = max.pixel.size)
      
      # plot with ggplot
      imgggplot <- ggplot2::ggplot(data.frame(x = 0, y = 0), 
                                   ggplot2::aes_string("x", "y")) + 
        ggplot2::coord_fixed(expand = FALSE, 
                             xlim = c(0, dim(img_raster)[2]), 
                             ylim = c(0, dim(img_raster)[1])) + 
        ggplot2::annotation_raster(img_raster, 0, dim(img_raster)[2], 
                                   dim(img_raster)[1], 0, interpolate = FALSE)
      imgggplot
    })
  }
  
  # Run the app
  shinyApp(ui = ui, server = server)
}
```

# Session info

```{r sessionInfo, echo=FALSE}
sessionInfo()
```
