
---
title: "`ImageArray`"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
package: "`r BiocStyle::pkg_ver('ImageArray')`"
author: 
  - name: Artur Manukyan
output:
  BiocStyle::html_document
vignette: |
  %\VignetteIndexEntry{ImageArray}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r load-libs}
library(BiocFileCache)
library(EBImage)
library(ImageArray)
library(ggplot2)
library(shiny)
```

# Introduction

`r Biocpkg("ImageArray")` is a package for Pyramidal and non-pyramidal 
images in `r Biocpkg("DelayedArray")` format.

## Installation

You can install `r Biocpkg("ImageArray")` from Bioconductor with:

```{r install-BiocManager, eval = FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE)) {
    install.packages("BiocManager")
}
BiocManager::install("ImageArray")
```

# Usage

The main usage of the `r Biocpkg("ImageArray")` is to deliver 
`r Biocpkg("DelayedArray")` operations for images or a list of images in 
Pyramidal format. You can also use
`magick-image` objects.

```{r read}
# make random EBImage image
f = system.file("images", "sample.png", package="EBImage")
img = readImage(f)

# create image array
dir.create(td <- tempfile())
output_h5 <- file.path(td, "h5test")
imgarray <- writeImgArray(img, 
                         format = "HDF5ImgArray", 
                         output = output_h5, 
                         replace = TRUE)
```

```{r visualize_read}
# visualize 
bfa.raster <- as.raster(imgarray, max.pixel.size = 300)
plot(bfa.raster)
```

Operations such as rotate, flip, flop or negate will be 
conducted on all levels

```{r rotate}
imgarray_rotated <- rotate(imgarray, degrees = 90)
imgarray_rotated

# visualize 
bfa.raster <- as.raster(imgarray_rotated, max.pixel.size = 300)
plot(bfa.raster)
```

You can even crop images, either by using `crop` or `[` method.

```{r crop}
imgarray_cropped <- imgarray[300:500, 200:300]
imgarray_cropped

# visualize 
bfa.raster <- as.raster(imgarray_cropped, max.pixel.size = 300)
plot(bfa.raster)
```

```{r crop2}
imgarray_cropped <- crop(imgarray, ind = list(300:500, 200:300))
imgarray_cropped

# visualize 
bfa.raster <- as.raster(imgarray_cropped, max.pixel.size = 300)
plot(bfa.raster)
```

# Using magick 

```{r read3}
# make random EBImage image
f = system.file("images", "sample.png", package="EBImage")
img = magick::image_read(f)

# create image array
dir.create(td <- tempfile())
output_h5 <- file.path(td, "h5test")
imgarray <- writeImgArray(img, 
                         format = "HDF5ImgArray", 
                         output = output_h5, 
                         replace = TRUE)

bfa.raster <- as.raster(imgarray, max.pixel.size = 300)
plot(bfa.raster)
```

# Using OME.TIFF

You can create `ImgArray` objects from OME.TIFF files with already defined
layers. 

```{r ometiff}
ome.tiff.file <- system.file("extdata", "xy_12bit__plant.ome.tiff", 
                             package = "ImageArray")
imgarray <- createImgArray(ome.tiff.file, series = 1, resolution = 1:2)
imgarray
```

Depending on the requested pixel size, functions may return the desired
resolution from the pyramid.

```{r ometiffvis, out.width="50%"}
bfa.raster <- as.raster(imgarray)
plot(bfa.raster)
dim(bfa.raster)
```

You can use either `max.pixel.size` or `min.pixel.size` to control the output

```{r ometiffvis2, out.width="50%"}
bfa.raster <- as.raster(img, max.pixel.size = 300)
plot(bfa.raster)
dim(bfa.raster)
```
# Using with Shiny

The pyramid scheme introduced by `ImgArray` objects can also be used to 
generate scalable plots and interactive Shiny application for visualizing 
large images. 

For this example, we will use a large H&E image used by 10x Genomics, 
generated after a Xenium in Situ platform run. 

```{r he}
image_file <- paste(
    "https://cf.10xgenomics.com/samples/xenium/1.0.1",
    "Xenium_FFPE_Human_Breast_Cancer_Rep1", 
    "Xenium_FFPE_Human_Breast_Cancer_Rep1_he_image.ome.tif",
    sep="/")
library(BiocFileCache)
bfc <- BiocFileCache()
image_file <- bfcrpath(bfc, image_file)
```

Let us create an `ImgArray` object first with the image above. The ome.tiff 
file includes multiple resolutions that we can ask `ImgArray` object to include.

```{r he_imgarray}
imgarray <- createImgArray(image_file, 
                           series = 1, 
                           resolution = 1:6)
imgarray
```
You can visualize subsets of large images really quickly using `crop` and
`as.raster`. Again, here we use `max.pixel.size` to regularize the resolution
parsed from the pyramid; that is, as defined below, only resolutions with
max pixel size 800 would be returned.

```{r speedy_vis, out.width="70%"}
# crop image
imgarray_sub <- crop(imgarray, ind = list(16000:19000, 7000:10000))

# convert to raster
img_raster <- as.raster(imgarray_sub, max.pixel.size = 800)

# plot with ggplot
imgggplot <- ggplot2::ggplot(data.frame(x = 0, y = 0), 
                             ggplot2::aes_string("x", "y")) + 
  ggplot2::coord_fixed(expand = FALSE, 
                       xlim = c(0, dim(img_raster)[2]), 
                       ylim = c(0, dim(img_raster)[1])) + 
  ggplot2::annotation_raster(img_raster, 0, dim(img_raster)[2], 
                             dim(img_raster)[1], 0, interpolate = FALSE)
imgggplot
```

Now let us create a shiny application where we can interactively define 
subsetting and visualize the image quickly. 

The Shiny application will react quickly to regardless of the size of the 
image subset. This is because `as.raster` determines the layer that has the
highest resolution without exceeding the threshold (`max.pixel.size`) 
automatically.

```{r speedy_vis_shiny}
if(interactive()){
  
  # variables
  dimimg <- dim(imgarray)
  max.pixel.size <- 800
  
  # Define UI
  ui <- fluidPage(
    sliderInput("x_slider", label = "X", min = 1, 
                max = dimimg[1], value = c(16000,19000)),
    sliderInput("y_slider", label = "Y", min = 1, 
                max = dimimg[2], value = c(7000,10000)),
    mainPanel(
      plotOutput(outputId = "scatterPlot")
    )
  )
  
  # Define server logic
  server <- function(input, output) {
    
    output$scatterPlot <- renderPlot({
      
      # crop image
      indx <- seq(input$x_slider[1], input$x_slider[2])
      indy <- seq(input$y_slider[1], input$y_slider[2])
      imgarray_sub <- crop(imgarray, ind = list(indx, indy))
      
      # convert to raster
      img_raster <- as.raster(imgarray_sub, max.pixel.size = max.pixel.size)
      
      # plot with ggplot
      imgggplot <- ggplot2::ggplot(data.frame(x = 0, y = 0), 
                                   ggplot2::aes_string("x", "y")) + 
        ggplot2::coord_fixed(expand = FALSE, 
                             xlim = c(0, dim(img_raster)[2]), 
                             ylim = c(0, dim(img_raster)[1])) + 
        ggplot2::annotation_raster(img_raster, 0, dim(img_raster)[2], 
                                   dim(img_raster)[1], 0, interpolate = FALSE)
      imgggplot
    })
  }
  
  # Run the app
  shinyApp(ui = ui, server = server)
}
```

# Session info

```{r sessionInfo, echo=FALSE}
sessionInfo()
```
