
---
title: "`ImageArray`"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
package: "`r BiocStyle::pkg_ver('ImageArray')`"
author: 
  - name: Artur Manukyan
output:
  BiocStyle::html_document
vignette: |
  %\VignetteIndexEntry{ImageArray}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=FALSE, message=FALSE, warning=FALSE)
```

```{r load-libs}
library(ImageArray)
```

# Introduction

ImageArray is a package for Pyramidal and non-pyramidal 
images in DelayedArray format.

# Usage

The main usage of the ImageArray is to deliver DelayedArray 
operations for images or a list of images in Pyramidal format.

```{r read}
# make random magick image
mat <- array(data=sample(1:255, 2000*5000*3, replace = TRUE), 
             dim=c(3, 2000, 5000))
mat_raster <- as.raster(aperm(mat, perm = c(2,3,1)), max = 255)
mat_image <- magick::image_read(mat_raster)

# create image array
dir.create(td <- tempfile())
output_h5 <- file.path(td, "h5test")
mat_list <- writeImgArray(mat_image, 
                          format = "HDF5ImgArray", 
                          output = output_h5, 
                          replace = TRUE)
mat_list
```

Operations such as rotate, flip, flop or negate will be 
conducted on all series

```{r rotate}
mat_list_rotated <- rotate(mat_list, degrees = 90)
mat_list_rotated
```

You can even crop images, either by using `crop` or `[` method.

```{r crop}
mat_list_cropped <- mat_list[2001:3000, 501:1000]
mat_list_cropped
```

```{r crop2}
mat_list_cropped <- crop(mat_list, ind = list(2001:3000, 501:1000))
mat_list_cropped
```

# Using OME.TIFF

You can create `ImgArray` objects from OME.TIFF files with already defined
layers. 

```{r ometiff}
ome.tiff.file <- system.file("extdata", "xy_12bit__plant.ome.tiff", 
                             package = "ImageArray")
img <- createImgArray(ome.tiff.file, n.series = 1, resolution = 1:2)
img
```

Depending on the requested pixel size, functions may return the desired
resolution from the pyramid.

```{r ometiffvis, out.width="50%"}
bfa.raster <- as.raster(img)
plot(bfa.raster)
dim(bfa.raster)
```

You can use either `max.pixel.size` or `min.pixel.size` to control the output

```{r ometiffvis2, out.width="50%"}
bfa.raster <- as.raster(img, max.pixel.size = 300)
plot(bfa.raster)
dim(bfa.raster)
```

# Session info

```{r sessionInfo, echo=FALSE}
sessionInfo()
```
